---
title: "Exploring the BRFSS data"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup

### Load packages

```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

Make sure your data and R Markdown files are in the same directory. When loaded
your data file will be called `brfss2013`. Delete this note when before you submit 
your work. 

```{r load-data}
load("brfss2013.RData")
```



* * *

## Part 1: Data
The Survey was initiated in 1984, with 15 states collecting surveillance data on risk behaviors through monthly telephone interviews. Over time, the number of states participating in the survey increased; by 2001, 50 states, the District of Columbia, Puerto Rico, Guam, and the US Virgin Islands were participating in the BRFSS.
This  sample shows the willingness of people o anwer the phone call and support in such long survey which makes the case a non-response bias; 

However it is not a huge representation of the general population, but the sample
size is useful to find statistical significance to even small differences.


For computing capabilities, we will refer to the data collected in the state
of California only. 
* * *

## Part 2: Research questions

**Research quesion 1:**
*What are the statistical features of the 
estimated age-gender specific maximum oxygen consumption? Does it follow a
normal distribution?*
For this e will use following variable:`$maxvo2_` for continuous in m1/min/kg

**Research quesion 2:** *Is there any correlation exist between the hours slept 
each night and reported days with poor physical health or poor mental health
in the last 30 days?*
Follwing variables is used:
`sleptim1` as explanatory variable which is in hours.
`physhlth` and `menthlth` as response variables which is in days.


**Research quesion 3:**
*Is there any correlation  between the body mass index
(BMI) of the individual answering the survey and having a high blood
cholesterol or having a heart attack?*
Following variable being used:
`_bmindex1` which being used as explanatory variables and is continuous.

`abcde2` and `fghi3` being used as responsive variables and is discrete.

* * *

## Part 3: Exploratory data analysis

NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research quesion 1:**
*What are the statistical features of the 
estimated age-gender specific maximum oxygen consumption? Does it follow a
normal distribution?*

The last two digits of this variable are implied decimal places, before doing 
anything this corrects to the units: `ml/min/kg`

```{r maxvo2.1}
df$maxvo2_ = as.numeric(df$maxvo2_)/100
```



**Research quesion 2:**

```{r}

```



**Research quesion 3:**

```{r}

```

